---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# slide

<!-- badges: start -->
[![Travis build status](https://travis-ci.org/DavisVaughan/slide.svg?branch=master)](https://travis-ci.org/DavisVaughan/slide)
[![Codecov test coverage](https://codecov.io/gh/DavisVaughan/slide/branch/master/graph/badge.svg)](https://codecov.io/gh/DavisVaughan/slide?branch=master)
<!-- badges: end -->

slide provides a family of general purpose "sliding window" functions. The API is purposefully _very_ similar to purrr, with functions such as `slide()`, `slide_dbl()`, `slide2()` and `pslide()`. The goal of these functions is usually to compute rolling averages, cumulative sums, rolling regressions, or other "window" based computations.

There are 3 reasons to use slide:

- Like `purrr::map()`, `slide()` is type stable, and also always returns a result with the same size as its input.

- Unlike `map()`, with data frames `slide()` iterates _row wise_. This is consistent with the theory that backs `slide()`, but also makes it a generic row wise data frame iterator that solves a number of problems outside of sliding windows. It just happens to be a neat side effect of this API.

- If you have ever needed to compute a rolling calculation _relative to an index_, then you might like `slide_index()`. This solves the problem of computing a "rolling mean over a 3 month window", where the number of days in each month is irregular.

If you are new to slide, but are familiar with purrr, I would encourage you to start with the documentation and examples for [`?slide`](https://davisvaughan.github.io/slide/reference/slide.html) and [`?slide_index`](https://davisvaughan.github.io/slide/reference/slide_index.html).

## Installation

slide is NOT yet on [CRAN](https://CRAN.R-project.org).

You can install the development version from [GitHub](https://github.com/) with:

``` r
remotes::install_github("DavisVaughan/slide")
```

(Just a warning that this uses a custom version of vctrs, which will soon become its development version).

## Examples

The [help page for `slide()`](https://davisvaughan.github.io/slide/reference/slide.html) has many examples, but here are a few:

```{r}
library(slide)
```

The classic example would be to do a moving average. `slide()` handles this with a combination of the `.before` and `.after` arguments, which control the width of the window and the alignment.

```{r}
# Moving average (Aligned right)
# "The current element + 2 elements before"
slide_dbl(1:5, ~mean(.x), .before = 2)

# Align left
# "The current element + 2 elements after"
slide_dbl(1:5, ~mean(.x), .after = 2)

# Center aligned
# "The current element + 1 element before + 1 element after"
slide_dbl(1:5, ~mean(.x), .before = 1, .after = 1)
```

With `unbounded()`, you can do a "cumulative slide" to compute cumulative expressions. I think of this as saying "give me everything before the current element."

```{r}
slide(1:4, ~.x, .before = unbounded())
```

With `.complete`, you can decide whether or not `.f` should be evaluated on incomplete windows. In the following example, the requested window size is 3, but the first two results are computed on windows of size 1 and 2 because partial results are allowed by default. When `.complete` is set to `TRUE`, the first two results are not computed.

```{r}
slide(1:4, ~.x, .before = 2)

slide(1:4, ~.x, .before = 2, .complete = TRUE)
```

## Data frames

Unlike `purrr::map()`, `slide()` iterates over data frames in a row wise fashion. Interestingly this means the default of `slide()` becomes a generic row wise iterator, with nice syntax for accessing data frame columns.

```{r}
cars <- mtcars[1:4,]

slide(cars, ~.x)

slide_dbl(cars, ~.x$mpg + .x$drat)
```

This makes rolling regressions trivial!

```{r, message=FALSE}
library(tibble)
set.seed(123)

df <- tibble(
  y = rnorm(100),
  x = rnorm(100)
)

# Window size of 20 rows
# The current row + 19 before
# (see slide_index() for how to do this relative to a date vector!)
df$regressions <- slide(df, ~lm(y ~ x, data = .x), .before = 19, .complete = TRUE)

df[15:25,]
```

## Index sliding

In many business settings, the value you want to compute is tied to some _index_, like a date vector. In these cases, you'll probably want to compute sliding windows relative to the index, and not using the fixed window that `slide()` provides. You can use `slide_index()` to pass in both `.x` and an index, `.i`, and the window will be calculated relative to that index. 

Here, when computing a "2 day window", you probably don't want `"2019-08-16"` and `"2019-08-18"` to be grouped together. `slide()` has no concept of an index, so when you specify a window size of 2, it will group these two together. `slide_index()`, on the other hand, will do the right thing.

```{r}
x <- 1:3
i <- as.Date(c("2019-08-15", "2019-08-16", "2019-08-18"))

# slide() has no concept of an "index"
slide(x, ~.x, .before = 1)

# "index aware"
slide_index(x, i, ~.x, .before = 1)
```

Essentially what happens is that when we get to `"2019-08-18"`, it "looks backwards" 1 day to set a window boundary at `"2019-08-17"`. Since the date at position 2, `"2019-08-16"`, is before `"2019-08-17"`, it is not included.

Powerfully, you can pass through any object to `.before` that computes a value from `.i - .before`. This means that you could also have used a lubridate period object (which gets even more interesting when you use `weeks()` or `months()`):

```{r}
slide_index(x, i, ~.x, .before = lubridate::days(1))
```

## Inspiration

This package is inspired heavily by SQL's window functions. The API is similar, but more general because you can iterate over any kind of R object.

There have been multiple attempts at creating sliding window functions (I personally created `rollify()`, and worked a little bit on `tsibble::slide()` with [Earo Wang](https://github.com/earowang)).

- `zoo::rollapply()`
- `tibbletime::rollify()`
- `tsibble::slide()`

I believe that slide is the next iteration of these. There are a few reasons for this:

- To me, the API is more intuitive, and is more flexible because `.before` and `.after` let you completely control the entry point (as opposed to fixed entry points like `"center"`, `"left"`, etc.

- It is objectively faster because it is written purely in C.

- With `slide_vec()` you can return any kind of object, and are not limited to the suffixed versions: `_dbl`, `_int`, etc.

- It iterates rowwise over data frames, consistent with the vctrs framework.

- I believe it is overall more consistent, backed by a theory that can always justify the sliding window generated by any combination of the parameters.

Earo and I have spoken, and we have mututally agreed that it would be best to deprecate `tsibble::slide()` in favor of `slide::slide()`.

Additionally, [data.table](https://github.com/Rdatatable/data.table)'s non-equi joins have been pretty much the only solution to the problem that `slide_index()` tries to solve. Their solution is robust and quite fast, and has been a nice benchmark for slide. slide is trying to solve a much narrower problem, so the API here is more focused.

## Performance

In terms of performance, be aware that any specialized package that shifts the function calls to C are going to be faster than slide. For example, `RcppRoll::roll_mean()` computes the rolling mean _at the C level_, which is bound to be faster. The purpose of slide is to be _general purpose_, while still being as fast as possible. This means that it can be used for more abstract things, like rolling regressions, or any other custom function that you want to use in a rolling fashion.

Otherwise, like `purrr::map()`, `slide()` is optimized in C to be as fast as possible, getting out of the way as quickly as it can so the main overhead are the `.f` calls.

## References

I've found the following references very useful to understand more about window functions:

- [Postgres SQL documentation](https://www.postgresql.org/docs/9.1/sql-expressions.html#SYNTAX-WINDOW-FUNCTIONS)

- [dbplyr window function vignette](https://dbplyr.tidyverse.org/articles/translation-function.html#window-functions)

- [SQLite documentation - with a flowchart](https://www.sqlite.org/windowfunctions.html)

- [Vertica Rows vs Range discussion](https://www.vertica.com/docs/9.2.x/HTML/Content/Authoring/SQLReferenceManual/Functions/Analytic/window_frame_clause.htm?origin_team=T02V9CHFH#ROWSversusRANGE)
